\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-10-18 k-planar.tex
recompiled the bibliography

The starting point for this work is the following `product structure theorem' for planar graphs\footnote{In this paper, all graphs are finite and undirected. Unless mentioned otherwise, all graphs are also simple. For any graph $G$ and any set $S$ (typically $S\subseteq V(G)$), let $G[S]$  denote the graph with vertex set $V(G)\cap S$ and edge set $\{uv\in E(G) : u,v\in S\}$.  We use $G-S$ as a shorthand for $G[V(G)\setminus S]$. Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20} (with an improvement by \citet{UWY}). A graph $G$ is \defin{contained} in a graph $X$ if $G$ is isomorphic to a subgraph of $X$.

\begin{thm}[\citep{DJMMUW20,UWY}]
\label{PlanarProduct}
Every planar graph is contained in:
\begin{compactenum}[(a)]
	\item $H\boxtimes P$ for some graph $H$ of treewidth at most $6$ and for some path $P$,
	\item $H\boxtimes P \boxtimes K_3$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
\end{compactenum}
\end{thm}

@@

\begin{table}[h]
  \centering{
    \begin{tabular}{lccl}
    	\hline
      {Graph class} & $\tw(H)$ & $\ell$ & {Reference} \\ \hline
      planar & $3^*$ & $3$ & \cite{DJMMUW20} \\
      planar & $4^*$ & $2$ & \cite{bose.morin.ea:optimal} \\
      planar & $6^*$ & $1$ & \cite{UWY} \\
      genus $g$ & $3^*$ & $\max\{2g,3\}$ & \cite{DHHW} \\
      genus $g$ & $2g+6$ & 1 & \cite{UWY} \\
      apex-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
      % B-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
      $k$-planar & $\binom{k+4}{3}-1$ & $18k^2 + 48k + 30$ & \cref{kPlanarProduct} \\
      $(g,k)$-planar & $\binom{k+4}{3}-1$ & $\max\{2g,3\}(6k^2+16k+10)$ & \cref{gkPlanarProduct} \\
      $(g,\delta)$-string & $\binom{\delta+4}{3}-1$ & $\max\{2g,3\}(\delta^4 + 4\delta^3 + 9\delta^2 + 10\delta+4)$ & \cref{StringProduct} \\
      $k$-nearest-neighbour & $O(k^6)$ & $O(k^4)$ & \cref{nn_product_structure} \\
      $d$-framed & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{d_framed_product_stucture} \\
      $1$-planar & $3^*$ & $7$ & \cref{1_planar_product} \\
      $d$-map & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{dMapProduct} \\
      $(g,d)$-map & $9$ & $\max\{2g,3\}\,(7d^2 -21d)$ & \cref{gdMapProduct} \\ \hline
      \multicolumn{4}{l}{$^*$ these bounds also apply to the simple treewidth of $H$.}
    \end{tabular}
  }
  \caption{Product structure theorems (of the form $G\subseteq H\boxtimes P\boxtimes K_\ell$).}
  \label{summary_table}
\end{table}

@@

As an example of the use of layered partitions, to prove \cref{PlanarProduct}(a),
\citet{DJMMUW20} build on an earlier result of \citet{PS21} to show that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$ (improved to $6$ in \cite{UWY}). The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering.\footnote{\citet{bose.morin.ea:optimal} have recently given linear time algorithms for computing layered $H$-partitions of planar graphs.}

@@

\begin{thm}[\citep{DJMMUW20,UWY,DHHW}]\label{GenusProduct}
  Every graph of Euler genus $g$ is contained in:
  \begin{compactenum}[(a)]
  \item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
  \item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some planar graph $H$ of treewidth at most $3$ and for some path $P$.
  \end{compactenum}
\end{thm}

@@

%\bibliographystyle{DavidNatbibStyle}
%\bibliography{k-planar}

\begin{thebibliography}{49}
	\providecommand{\natexlab}[1]{#1}
	\providecommand{\msn}[1]{MR:\,\href{http://www.ams.org/mathscinet-getitem?mr=MR{#1}}{#1}}
	\providecommand{\ZBL}[1]{Zbl:\,\href{https://www.zentralblatt-math.org/zmath/en/search/?q=an:#1}{#1}}
	\providecommand{\url}[1]{\texttt{#1}}
	\providecommand{\urlprefix}{}
	\expandafter\ifx\csname urlstyle\endcsname\relax
	\providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
	\providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
		\urlstyle{rm}\Url}\fi

	\bibitem[{{\'{A}}brego et~al.(2011){\'{A}}brego, Monroy,
		Fern{\'{a}}ndez{-}Merchant, Flores{-}Pe{\~{n}}aloza, Hurtado,
		Sacrist{\'{a}}n, and Saumell}]{abrego.munroy.ea:on}
	\textsc{Bernardo~M. {\'{A}}brego, Ruy~Fabila Monroy, Silvia
		Fern{\'{a}}ndez{-}Merchant, David Flores{-}Pe{\~{n}}aloza, Ferran Hurtado,
		Vera Sacrist{\'{a}}n, and Maria Saumell}.
	\newblock \href{https://doi.org/10.1016/j.comgeo.2010.11.003}{On crossing
		numbers of geometric proximity graphs}.
	\newblock \emph{Comput. Geom.}, 44(4):216--233, 2011.

	\bibitem[{Aigner and Ziegler(2010)}]{Proofs4}
	\textsc{Martin Aigner and G{\"u}nter~M. Ziegler}.
	\newblock Proofs from {T}he {B}ook.
	\newblock Springer, 4th edn., 2010.

	\bibitem[{Ajtai et~al.(1982)Ajtai, Chv{\'a}tal, Newborn, and
		Szemer{\'e}di}]{ajtai.chvatal.ea:crossing-free}
	\textsc{Mikl{\'o}s Ajtai, Va{\v{s}}ek Chv{\'a}tal, Monroe~M. Newborn, and Endre
		Szemer{\'e}di}.
	\newblock \href{https://doi.org/10.1016/S0304-0208(08)73484-4}{Crossing-free
		subgraphs}.
	\newblock \emph{Ann. Discrete Math.}, 12:9--12, 1982.

	\bibitem[{Alam et~al.(2020)Alam, Bekos, Gronemann, Kaufmann, and
		Pupyrev}]{ABGKP20}
	\textsc{Jawaherul~Md. Alam, Michael~A. Bekos, Martin Gronemann, Michael
		Kaufmann, and Sergey Pupyrev}.
	\newblock \href{https://doi.org/10.1007/s00453-020-00697-4}{Queue layouts of
		planar 3-trees}.
	\newblock \emph{Algorithmica}, 82:2564--2585, 2020.

	\bibitem[{Alon et~al.(2002)Alon, Grytczuk, Ha{\l}uszczak, and
		Riordan}]{AGHR-RSA02}
	\textsc{Noga Alon, {Jaros{\l}aw} Grytczuk, Mariusz Ha{\l}uszczak, and Oliver
		Riordan}.
	\newblock \href{https://doi.org/10.1002/rsa.10057}{Nonrepetitive colorings of
		graphs}.
	\newblock \emph{Random Structures Algorithms}, 21(3-4):336--346, 2002.

	\bibitem[{Babai et~al.(1982)Babai, Chung, Erd\H{o}s, Graham, and
		Spencer}]{babai.chung.ea:on}
	\textsc{Laszlo Babai, F.~R.~K. Chung, Paul Erd\H{o}s, Ronald~L. Graham, and
		J.~H. Spencer}.
	\newblock On graphs which contain all sparse graphs.
	\newblock In \emph{Theory and practice of combinatorics. A collection of
		articles honoring Anton Kotzig on the occasion of his sixtieth birthday}, pp.
	21--26. Elsevier, 1982.

	\bibitem[{Bekos et~al.(2020)Bekos, Lozzo, Griesbach, Gronemann, Montecchiani,
		and Raftopoulou}]{BDGGMR}
	\textsc{Michael~A. Bekos, Giordano~Da Lozzo, Svenja Griesbach, Martin
		Gronemann, Fabrizio Montecchiani, and Chrysanthi~N. Raftopoulou}.
	\newblock \href{https://doi.org/10.4230/LIPIcs.SoCG.2020.16}{Book embeddings of
		nonplanar graphs with small faces in few pages}.
	\newblock In \textsc{Sergio Cabello and Danny~Z. Chen}, eds., \emph{36th
		International Symposium on Computational Geometry, SoCG 2020, June 23-26,
		2020, Z{\"{u}}rich, Switzerland}, vol. 164 of \emph{LIPIcs}, pp. 16:1--16:17.
	Schloss Dagstuhl - Leibniz-Zentrum f{\"{u}}r Informatik, 2020.
	\newblock arXiv:2003.07655.

	\bibitem[{Bodlaender(1998)}]{Bodlaender-TCS98}
	\textsc{Hans~L. Bodlaender}.
	\newblock \href{https://doi.org/10.1016/S0304-3975(97)00228-4}{A partial
		$k$-arboretum of graphs with bounded treewidth}.
	\newblock \emph{Theoret. Comput. Sci.}, 209(1-2):1--45, 1998.

	\bibitem[{Bonnet et~al.(2022)Bonnet, Kwon, and Wood}]{BKW}
	\textsc{\'Edouard Bonnet, {O-joung} Kwon, and David~R. Wood}.
	\newblock \href{http://arxiv.org/abs/2202.11858}{Reduced bandwidth: a
		qualitative strengthening of twin-width in minor-closed classes (and
		beyond)}.
	\newblock 2022, arXiv:2202.11858.

	\bibitem[{Bose et~al.(2012)Bose, Dujmovic, Hurtado, Iacono, Langerman, Meijer,
		Adinolfi, Saumell, and Wood}]{ProximityGraphs}
	\textsc{Prosenjit Bose, Vida Dujmovic, Ferran Hurtado, John Iacono, Stefan
		Langerman, Henk Meijer, Vera~Sacrist{\'{a}}n Adinolfi, Maria Saumell, and
		David~R. Wood}.
	\newblock \href{https://doi.org/10.1142/S0218195912500112}{Proximity graphs: E,
		{\(\delta\)}, {\(\Delta\)}, {\(\chi\)} and {\(\omega\)}}.
	\newblock \emph{Int. J. Comput. Geometry Appl.}, 22(5):439--470, 2012.

	\bibitem[{Bose et~al.(2020)Bose, Dujmovi\'c, Javarsineh, and Morin}]{BDJM}
	\textsc{Prosenjit Bose, Vida Dujmovi\'c, Mehrnoosh Javarsineh, and Pat Morin}.
	\newblock \href{http://arxiv.org/abs/2007.06455}{Asymptotically optimal vertex
		ranking of planar graphs}.
	\newblock 2020, arXiv:2007.06455.

	\bibitem[{Bose et~al.(2022)Bose, Morin, and Odak}]{bose.morin.ea:optimal}
	\textsc{Prosenjit Bose, Pat Morin, and Saeed Odak}.
	\newblock \href{http://arxiv.org/abs/2202.08870}{An optimal algorithm for
		product structure in planar graphs}.
	\newblock 2022, arXiv:2202.08870.

	\bibitem[{Bose et~al.(2001)Bose, Morin, Stojmenović, and
		Urrutia}]{bose.morin.ea:routing}
	\textsc{Prosenjit Bose, Pat Morin, Ivan Stojmenović, and Jorge Urrutia}.
	\newblock \href{https://doi.org/10.1023/A:1012319418150}{Routing with
		guaranteed delivery in ad hoc wireless networks}.
	\newblock \emph{Wireless Networks}, 7(6):609--616, 2001.

	\bibitem[{Brandenburg(2019)}]{Brandenburg19}
	\textsc{Franz~J. Brandenburg}.
	\newblock \href{https://doi.org/10.1007/s00453-018-0510-x}{Characterizing and
		recognizing 4-map graphs}.
	\newblock \emph{Algorithmica}, 81(5):1818--1843, 2019.

	\bibitem[{Brandenburg(2020)}]{Brandenburg20}
	\textsc{Franz~J. Brandenburg}.
	\newblock \href{http://arxiv.org/abs/2012.06874}{Book embeddings of $k$-map
		graphs}.
	\newblock 2020, arXiv:2012.06874.

	\bibitem[{Chen et~al.(2002)Chen, Grigni, and Papadimitriou}]{CGP02}
	\textsc{Zhi-Zhong Chen, Michelangelo Grigni, and Christos~H. Papadimitriou}.
	\newblock \href{https://doi.org/10.1145/506147.506148}{Map graphs}.
	\newblock \emph{J. ACM}, 49(2):127--138, 2002.

	\bibitem[{Chen et~al.(2006)Chen, Grigni, and Papadimitriou}]{CGP06}
	\textsc{Zhi-Zhong Chen, Michelangelo Grigni, and Christos~H. Papadimitriou}.
	\newblock \href{https://doi.org/10.1007/s00453-005-1184-8}{Recognizing
		hole-free 4-map graphs in cubic time}.
	\newblock \emph{Algorithmica}, 45(2):227--262, 2006.

	\bibitem[{Diestel(2018)}]{Diestel5}
	\textsc{Reinhard Diestel}.
	\newblock \href{http://diestel-graph-theory.com/}{Graph theory}, vol. 173 of
	\emph{Graduate Texts in Mathematics}.
	\newblock Springer, 5th edn., 2018.

	\bibitem[{Distel et~al.(2021)Distel, Hickingbotham, Huynh, and Wood}]{DHHW}
	\textsc{Marc Distel, Robert Hickingbotham, Tony Huynh, and David~R. Wood}.
	\newblock \href{http://arxiv.org/abs/2112.10025}{Improved product structure for
		graphs on surfaces}.
	\newblock 2021, arXiv:2112.10025.

	\bibitem[{D\k{e}bski et~al.(2021)D\k{e}bski, Felsner, Micek, and
		Schr\"{o}der}]{DFMS21}
	\textsc{Micha{\l} D\k{e}bski, Stefan Felsner, Piotr Micek, and Felix
		Schr\"{o}der}.
	\newblock \href{https://doi.org/10.19086/aic.27351}{Improved bounds for
		centered colorings}.
	\newblock \emph{Adv. Comb.}, \#8, 2021.

	\bibitem[{Dujmovi\'c et~al.(2017{\natexlab{a}})Dujmovi\'c, Eppstein, and
		Wood}]{dujmovic.eppstein.ea:structure}
	\textsc{Vida Dujmovi\'c, David Eppstein, and David~R. Wood}.
	\newblock \href{https://doi.org/10.1137/16M1062879}{Structure of graphs with
		locally restricted crossings}.
	\newblock \emph{{SIAM} J. Discrete Math.}, 31(2):805--824, 2017{\natexlab{a}}.

	\bibitem[{Dujmovi\'c et~al.(2021)Dujmovi\'c, Esperet, Gavoille, Joret, Micek,
		and Morin}]{DEJGMM21}
	\textsc{Vida Dujmovi\'c, Louis Esperet, Cyril Gavoille, Gwena\"el Joret, Piotr
		Micek, and Pat Morin}.
	\newblock \href{https://doi.org/10.1145/3477542}{Adjacency labelling for planar
		graphs (and beyond)}.
	\newblock \emph{J. ACM}, 68(6):42, 2021.

	\bibitem[{Dujmovi\'c et~al.(2020)Dujmovi\'c, Esperet, Joret, Walczak, and
		Wood}]{dujmovic.esperet.ea:planar}
	\textsc{Vida Dujmovi\'c, Louis Esperet, Gwena{\"{e}}l Joret, Bartosz Walczak,
		and David~R. Wood}.
	\newblock \href{https://doi.org/10.19086/aic.12100}{Planar graphs have bounded
		nonrepetitive chromatic number}.
	\newblock \emph{Adv. Comb.}, \#5, 2020.

	\bibitem[{Dujmovi{\'c} et~al.(2022)Dujmovi{\'c}, Esperet, Morin, Walczak, and
		Wood}]{DEMWW22}
	\textsc{Vida Dujmovi{\'c}, Louis Esperet, Pat Morin, Bartosz Walczak, and
		David~R. Wood}.
	\newblock \href{https://doi.org/10.1017/S0963548321000213}{Clustered
		3-colouring graphs of bounded degree}.
	\newblock \emph{Combin. Probab. Comput.}, 31(1):123--135, 2022.

	\bibitem[{Dujmovi\'{c} et~al.(2016)Dujmovi\'{c}, Joret, Kozik, and
		Wood}]{DJKW16}
	\textsc{Vida Dujmovi\'{c}, Gwena\"{e}l Joret, Jakub Kozik, and David~R. Wood}.
	\newblock \href{https://doi.org/10.1007/s00493-015-3070-6}{Nonrepetitive
		colouring via entropy compression}.
	\newblock \emph{Combinatorica}, 36(6):661--686, 2016.

	\bibitem[{Dujmovi{\'c} et~al.(2020)Dujmovi{\'c}, Joret, Micek, Morin, Ueckerdt,
		and Wood}]{DJMMUW20}
	\textsc{Vida Dujmovi{\'c}, Gwena\"{e}l Joret, Piotr Micek, Pat Morin, Torsten
		Ueckerdt, and David~R. Wood}.
	\newblock \href{https://doi.org/10.1145/3385731}{Planar graphs have bounded
		queue-number}.
	\newblock \emph{J. ACM}, 67(4):\#22, 2020.

	\bibitem[{Dujmovi\'c et~al.(2022)Dujmovi\'c, Morin, and Odak}]{DMO}
	\textsc{Vida Dujmovi\'c, Pat Morin, and Saeed Odak}.
	\newblock \href{http://arxiv.org/abs/2202.12882}{Odd colourings of graph
		products}.
	\newblock 2022, arXiv:2202.12882.

	\bibitem[{Dujmovi\'c et~al.(2017{\natexlab{b}})Dujmovi\'c, Morin, and
		Wood}]{dujmovic.morin.ea:layered}
	\textsc{Vida Dujmovi\'c, Pat Morin, and David~R. Wood}.
	\newblock \href{https://doi.org/10.1016/j.jctb.2017.05.006}{Layered separators
		in minor-closed graph classes with applications}.
	\newblock \emph{J. Comb. Theory, Ser. {B}}, 127:111--147, 2017{\natexlab{b}}.

	\bibitem[{Dvo{\v{r}}{\'a}k et~al.(2021)Dvo{\v{r}}{\'a}k, Huynh, Joret, Liu, and
		Wood}]{DHJLW21}
	\textsc{Zden{\v{e}}k Dvo{\v{r}}{\'a}k, Tony Huynh, Gwena\"el Joret, Chun-Hung
		Liu, and David~R. Wood}.
	\newblock \href{https://doi.org/10.1007/978-3-030-62497-2_32}{Notes on graph
		product structure theory}.
	\newblock In \textsc{David~R. Wood, Jan de~Gier, Cheryl~E. Praeger, and Terence
		Tao}, eds., \emph{2019-20 MATRIX Annals}, pp. 513--533. Springer, 2021.
	\newblock arXiv:2001.08860.

	\bibitem[{Elmallah and Colbourn(1990)}]{elmallah.colbourn:on}
	\textsc{Ehab~S. Elmallah and Charles~J. Colbourn}.
	\newblock \href{https://doi.org/10.1016/0012-365X(90)90293-Q}{On two dual
		classes of planar graphs}.
	\newblock \emph{Discrete. Math.}, 80(1):21--40, 1990.

	\bibitem[{Esperet et~al.(2020)Esperet, Joret, and Morin}]{EJM}
	\textsc{Louis Esperet, Gwena\"{e}l Joret, and Pat Morin}.
	\newblock \href{http://arxiv.org/abs/2010.05779}{Sparse universal graphs for
		planarity}.
	\newblock 2020, arXiv:2010.05779.

	\bibitem[{Fomin et~al.(2012)Fomin, Lokshtanov, and Saurabh}]{FLS-SODA12}
	\textsc{Fedor~V. Fomin, Daniel Lokshtanov, and Saket Saurabh}.
	\newblock \href{https://doi.org/10.1137/1.9781611973099.124}{Bidimensionality
		and geometric graphs}.
	\newblock In \emph{Proc. 23rd {A}nnual {ACM}-{SIAM} {S}ymposium on {D}iscrete
		{A}lgorithms}, pp. 1563--1575. 2012.

	\bibitem[{Fox and Pach(2010)}]{FP10}
	\textsc{Jacob Fox and J{\'a}nos Pach}.
	\newblock \href{https://doi.org/10.1017/S0963548309990459}{A separator theorem
		for string graphs and its applications}.
	\newblock \emph{Combin. Probab. Comput.}, 19(3):371--390, 2010.

	\bibitem[{Fox and Pach(2014)}]{FP14}
	\textsc{Jacob Fox and J{\'a}nos Pach}.
	\newblock \href{https://doi.org/10.1017/S0963548313000412}{Applications of a
		new separator theorem for string graphs}.
	\newblock \emph{Combin. Probab. Comput.}, 23(1):66--74, 2014.

	\bibitem[{Harvey and Wood(2017)}]{HW17}
	\textsc{Daniel~J. Harvey and David~R. Wood}.
	\newblock \href{https://doi.org/10.1002/jgt.22030}{Parameters tied to
		treewidth}.
	\newblock \emph{J. Graph Theory}, 84(4):364--385, 2017.

	\bibitem[{Heath et~al.(1992)Heath, Leighton, and Rosenberg}]{HLR92}
	\textsc{Lenwood~S. Heath, F.~Thomson Leighton, and Arnold~L. Rosenberg}.
	\newblock \href{https://doi.org/10.1137/0405031}{Comparing queues and stacks as
		mechanisms for laying out graphs}.
	\newblock \emph{SIAM J. Discrete Math.}, 5(3):398--412, 1992.

\bibitem[{van~den Heuvel and Wood(2018)}]{vdHW18}
\textsc{Jan van~den Heuvel and David~R. Wood}.
\newblock \href{https://doi.org/10.1112/jlms.12127}{Improper colourings
	inspired by {H}adwiger's conjecture}.
\newblock \emph{J. London Math. Soc.}, 98:129--148, 2018.
\newblock arXiv:1704.06536.

	\bibitem[{Kannan et~al.(1992)Kannan, Naor, and
		Rudich}]{kannan.naor.ea:implicit}
	\textsc{Sampath Kannan, Moni Naor, and Steven Rudich}.
	\newblock \href{https://doi.org/10.1137/0405049}{Implicit representation of
		graphs}.
	\newblock \emph{{SIAM} J. Discrete Math.}, 5(4):596--603, 1992.

	\bibitem[{Knauer and Ueckerdt(2012)}]{knauer.ueckerdt:simple}
	\textsc{Kolja Knauer and Torsten Ueckerdt}.
	\newblock
	\href{https://kam.mff.cuni.cz/workshops/mcw/work18/mcw2012booklet.pdf}{Simple
		treewidth}.
	\newblock In \textsc{Pavel Rytír}, ed., \emph{Midsummer Combinatorial Workshop
		Prague}. 2012.

	\bibitem[{Kobourov et~al.(2017)Kobourov, Liotta, and
		Montecchiani}]{kobourov.liotta.ea:annotated}
	\textsc{Stephen~G. Kobourov, Giuseppe Liotta, and Fabrizio Montecchiani}.
	\newblock \href{https://doi.org/10.1016/j.cosrev.2017.06.002}{An annotated
		bibliography on 1-planarity}.
	\newblock \emph{Comput. Sci. Rev.}, 25:49--67, 2017.

	\bibitem[{Kratochv{\'{\i}}l and Vaner(2012)}]{kratochvil.vaner:note}
	\textsc{Jan Kratochv{\'{\i}}l and Michal Vaner}.
	\newblock \href{http://arxiv.org/abs/1210.8113}{A note on planar partial
		3-trees}.
	\newblock 2012, arXiv:1210.8113.

	\bibitem[{Leighton(1983)}]{leighton:complexity}
	\textsc{F.~Thomson Leighton}.
	\newblock Complexity issues in {VLSI}.
	\newblock MIT Press, Cambridge, MA, 1983.

	\bibitem[{Mohar and Thomassen(2001)}]{mohar.thomassen:graphs}
	\textsc{Bojan Mohar and Carsten Thomassen}.
	\newblock Graphs on surfaces.
	\newblock Johns Hopkins University Press, 2001.

	\bibitem[{Ne{\v{s}}et{\v{r}}il and Ossona~de Mendez(2012)}]{Sparsity}
	\textsc{Jaroslav Ne{\v{s}}et{\v{r}}il and Patrice Ossona~de Mendez}.
	\newblock \href{https://doi.org/10.1007/978-3-642-27875-4}{Sparsity}, vol.~28
	of \emph{Algorithms and Combinatorics}.
	\newblock Springer, 2012.

	\bibitem[{Pach and T{\'o}th(2002)}]{PachToth-DCG02}
	\textsc{J{\'a}nos Pach and G{\'e}za T{\'o}th}.
	\newblock \href{https://doi.org/10.1007/s00454-002-2891-4}{Recognizing string
		graphs is decidable}.
	\newblock \emph{Discrete Comput. Geom.}, 28(4):593--606, 2002.

	\bibitem[{Pilipczuk and Siebertz(2021)}]{PS21}
	\textsc{Micha{\l} Pilipczuk and Sebastian Siebertz}.
	\newblock \href{https://doi.org/10.1016/j.jctb.2021.06.002}{Polynomial bounds
		for centered colorings on proper minor-closed graph classes}.
	\newblock \emph{J. Comb. Theory, Series B}, 151:111--147, 2021.

	\bibitem[{Reed(2003)}]{Reed03}
	\textsc{Bruce~A. Reed}.
	\newblock \href{https://doi.org/10.1007/0-387-22444-0_4}{Algorithmic aspects of
		tree width}.
	\newblock In \emph{Recent advances in algorithms and combinatorics}, vol.~11,
	pp. 85--107. Springer, 2003.

	\bibitem[{Ueckerdt et~al.(2021)Ueckerdt, Wood, and Yi}]{UWY}
	\textsc{Torsten Ueckerdt, David~R. Wood, and Wendy Yi}.
	\newblock \href{https://doi.org/10.37236/10614}{An improved planar graph product structure theorem}.
	\newblock Electronic J. Combinatorics 29:P2.51, 2022.

	\bibitem[{Zhu(2009)}]{zhu:colouring}
	\textsc{Xuding Zhu}.
	\newblock \href{https://doi.org/10.1016/j.disc.2008.03.024}{Colouring graphs
		with bounded generalized colouring number}.
	\newblock \emph{Discrete Math.}, 309(18):5562--5568, 2009.

\end{thebibliography}

\end{document}

>>>>

The starting point for this work is the following `product structure theorem' for planar graphs\footnote{In this paper, all graphs are finite and undirected. Unless mentioned otherwise, all graphs are also simple. For any graph $G$ and any set $S$ (typically $S\subseteq V(G)$), let $G[S]$  denote the graph with vertex set $V(G)\cap S$ and edge set $\{uv\in E(G) : u,v\in S\}$.  We use $G-S$ as a shorthand for $G[V(G)\setminus S]$. Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20} (with an improvement by \citet{UWY22}). A graph $G$ is \defin{contained} in a graph $X$ if $G$ is isomorphic to a subgraph of $X$.

\begin{thm}[\citep{DJMMUW20,UWY22}]
\label{PlanarProduct}
Every planar graph is contained in:
\begin{compactenum}[(a)]
	\item $H\boxtimes P$ for some graph $H$ of treewidth at most $6$ and for some path $P$,
	\item $H\boxtimes P \boxtimes K_3$ for some graph $H$ of treewidth at most $3$ and for some path $P$.
\end{compactenum}
\end{thm}

@@

\begin{table}[h]
  \centering{
    \begin{tabular}{lccl}
    	\hline
      {Graph class} & $\tw(H)$ & $\ell$ & {Reference} \\ \hline
      planar & $3^*$ & $3$ & \cite{DJMMUW20} \\
      planar & $4^*$ & $2$ & \cite{bose.morin.ea:optimal} \\
      planar & $6^*$ & $1$ & \cite{UWY22} \\
      genus $g$ & $3^*$ & $\max\{2g,3\}$ & \cite{DHHW} \\
      genus $g$ & $2g+6$ & 1 & \cite{UWY22} \\
      apex-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
      % B-minor-free & $O(1)$ & $O(1)$ & \cite{DJMMUW20} \\
      $k$-planar & $\binom{k+4}{3}-1$ & $18k^2 + 48k + 30$ & \cref{kPlanarProduct} \\
      $(g,k)$-planar & $\binom{k+4}{3}-1$ & $\max\{2g,3\}(6k^2+16k+10)$ & \cref{gkPlanarProduct} \\
      $(g,\delta)$-string & $\binom{\delta+4}{3}-1$ & $\max\{2g,3\}(\delta^4 + 4\delta^3 + 9\delta^2 + 10\delta+4)$ & \cref{StringProduct} \\
      $k$-nearest-neighbour & $O(k^6)$ & $O(k^4)$ & \cref{nn_product_structure} \\
      $d$-framed & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{d_framed_product_stucture} \\
      $1$-planar & $3^*$ & $7$ & \cref{1_planar_product} \\
      $d$-map & $3^*$ & $d+ 3\floor{d/2} - 3$ & \cref{dMapProduct} \\
      $(g,d)$-map & $9$ & $\max\{2g,3\}\,(7d^2 -21d)$ & \cref{gdMapProduct} \\ \hline
      \multicolumn{4}{l}{$^*$ these bounds also apply to the simple treewidth of $H$.}
    \end{tabular}
  }
  \caption{Product structure theorems (of the form $G\subseteq H\boxtimes P\boxtimes K_\ell$).}
  \label{summary_table}
\end{table}

@@

As an example of the use of layered partitions, to prove \cref{PlanarProduct}(a),
\citet{DJMMUW20} build on an earlier result of \citet{PS21} to show that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$ (improved to $6$ in \citep{UWY22}). The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering.\footnote{\citet{bose.morin.ea:optimal} have recently given linear time algorithms for computing layered $H$-partitions of planar graphs.}

@@

\begin{thm}[\citep{DJMMUW20,UWY22,DHHW}]\label{GenusProduct}
  Every graph of Euler genus $g$ is contained in:
  \begin{compactenum}[(a)]
  \item $H  \boxtimes P$ for some graph $H$ of treewidth at most $2g+6$  and some path $P$.
  \item $H \boxtimes P \boxtimes K_{\max\{2g,3\}}$ for some planar graph $H$ of treewidth at most $3$ and for some path $P$.
  \end{compactenum}
\end{thm}

@@

\bibliographystyle{DavidNatbibStyle}
\bibliography{k-planar}
\end{document}
